---
title: "Predicting woody cover with enamel carbon isotopes using Bayesian multi-level modeling"
author: "Maddie Kelly"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages & custom functions
```{r}
library(tidyverse)
library(ggmap)
library(brms)
library(tidybayes)
library(bayesplot)
library(patchwork)
library(ggnewscale)
library(ggstar)
```

```{r}
# custom functions for plotting (color/fill scales)
source(file = "code/custom_functions.R", local = T)
```

# Import dataset
```{r}
iso_data <- read_csv("data/modern_enamel_d13C.csv")

# the d13C_corrected column contains the d13C values corrected to a pre-industrial atmospheric d13C, which are used for all analyses

# center and scale d13C values 
iso_data$d13C_scaled <- scale(iso_data$d13C_corrected)
```

## Map of sites (Figure 1A)
```{r mapping}
# use stadia maps
register_stadiamaps(key = "YOUR-KEY-HERE") # input personal Stadia maps API key to generate map
terrainmap <- get_map(location = c(left = -20, bottom = -37, right = 50, top = 23), zoom = 4, maptype = "stamen_terrain_background", source = "stadia")
finelines <- get_map(location = c(left = -20, bottom = -37, right = 50, top = 23), zoom = 4, maptype = "stamen_toner_lines", source = "stadia")

terrainmap_attributes <- attributes(terrainmap)
terrainmap_transparent <- matrix(adjustcolor(terrainmap, 
                                                alpha.f = 0.9), 
                                    nrow = nrow(terrainmap))
attributes(terrainmap_transparent) <- terrainmap_attributes

# plotting all of the localities included in the analysis 
isotope_data_map <- ggmap(terrainmap_transparent)+
  inset_ggmap(finelines)+ # sticks the country lines on top 
  geom_point(data = iso_data, aes(x = Longitude_updated, y = Latitude_updated), fill = "#A5506DFF", size = 3, shape = 24)+
  scale_y_continuous(position = "right", expand = 0)+
  labs(x = "Longitude", y = "Latitude", title = "Modern enamel carbon isotope data localities", size = 12)+
  annotate("text",label = "© Stadia Maps © Stamen Design\n© OpenMapTiles © OpenStreetMap contributors", x = -18, y = -33, size = 3, hjust=0)+
  theme(axis.text = element_text(size = 10), plot.title = element_text(hjust = 0.5, size = 14))

isotope_data_map
ggsave(isotope_data_map, filename = "figures/isotope_data_map.pdf")

```

## Barplot of sample size by family (Figure 1B):
```{r}
iso_data %>% group_by(Family) %>% tally() %>% 
  ggplot() +
  geom_col(aes(x = reorder(Family, n, decreasing = T), y = n, color = Family, fill = Family), alpha = 0.4, linewidth = 2) +
  family_colors()+
  family_fills()+
  labs(x = "Family", y = "Number of samples")+
  theme_classic(base_size = 18)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/d13C_family_plot.png", width = 9, height = 7)
```

# Model with simulated data

## 1. Simulate slope and intercept parameters for a group of 10, random hypothetical taxa. 

Three taxa will be given very strong slopes, three given mild slopes, and the remaining four given weak slopes, to simulate the variation in the relationship between d13C and woody cover in real mammal taxa. The intercepts will be drawn at random from a Gaussian distribution with a mean 0.5 and standard deviation 0.25. In this case, the slopes and intercepts are not explicitly modeled to covary. 
```{r}
ggplot()+
  geom_function(fun = dnorm, args = list(mean = -0.05, sd = 0.02), aes(color = "very strong"), linewidth = 2)+
  geom_function(fun = dnorm, args = list(mean = -0.2, sd = 0.05), aes(color = "mild"), linewidth = 2)+
  geom_function(fun = dnorm, args = list(mean = -0.5, sd = 0.1), aes(color = "weak"), linewidth = 2)+
  scale_color_manual(values = c("very strong" = "mediumpurple", "mild" = "coral2", "weak" = "darkgreen"), name = "varying slopes")+
  xlim(-1,0.5)+
  theme_classic()
ggsave("figures/simulation_slope_distributions.png")

set.seed(2840) # set seed for reproducibility
sim_taxa <- data.frame(taxon = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"))
sim_taxa <- sim_taxa %>% mutate(intercept = rnorm(n(), mean = 0.5, sd = 0.25)) # generate random intercept values for each taxon
sim_taxa$slope <- NA
sim_taxa[1:3,] <- sim_taxa[1:3,] %>% 
  mutate(slope = rnorm(n(), mean = -0.05, sd = 0.02)) # generate a set of very strong slopes
sim_taxa[4:6,] <- sim_taxa[4:6,] %>%
  mutate(slope = rnorm(n(), mean = -0.2, sd = 0.05)) # generate a set of weaker slopes
sim_taxa[7:10,] <- sim_taxa[7:10,] %>%
  mutate(slope = rnorm(n(), mean = -0.5, sd = 0.1)) # generate a set of very weak slopes
```

## 2. Simulate a set of sites (n = 20) with random woody cover values, and with 10 of each taxon at each site (n = 100 individuals per site). 

Using the taxon specific slope and intercept parameters, paired with the site's woody cover value, generate simulated d13C values for each individual at each site. 
```{r}
set.seed(2840) # for reproducibility
nsites = 20
mu = 0.35
phi = 7.5
dsim <- tibble(site = 1:nsites,
               wc_sim = rbeta(n = nsites, shape1 = mu*phi, shape2 = (1-mu)*phi)) # draw woody cover values from this Beta distribution; mu and phi (mean and variance) of the beta distribution are converted to shape1 and shape2 parameters using the equations shape1 = mu * phi and shape2 = (1-mu)*phi.

# repeat rows of simulated woody cover dataset to generate multiple observations for each "site". 
dsim <- dsim %>% slice(rep(1:n(), each = 10))
dsim$site <- as.factor(dsim$site)

# simulate "carbon isotope" values for each hypothetical taxon, for each site, using the woody cover value and intercept and beta parameters set for each taxon
for (i in 1:nrow(sim_taxa)) {
  for (j in 1:nrow(dsim)) { 
    dsim[j,paste("d13C_sim_", sim_taxa[i,"taxon"], sep = "")] <- rnorm(1, mean = (dsim[j,]$wc_sim - sim_taxa[i,"intercept"]) / sim_taxa[i, "slope"], sd = 1.5)
  }
}

# does that generate the relationships I want to see?
plot(density(dsim$d13C_sim_A))
plot(dsim$wc_sim~dsim$d13C_sim_A)

# pivot the data so there's only one d13C column and another column indicating taxon
dsim <- dsim %>% pivot_longer(d13C_sim_A:d13C_sim_J)

# rename columns and column values for taxa
colnames(dsim)[colnames(dsim) %in% c("name", "value")] <- c("taxon", "d13C_sim")
dsim$taxon <- sub("d13C_sim_", "", dsim$taxon)

# scale d13C values 
dsim$d13C_scaled <- scale(dsim$d13C_sim)

```

## 3. Visualize these simulated relationships to make sure that the data generation process generated the expected results.
```{r}
ggplot(data = dsim, aes(x = d13C_sim, y = wc_sim, color = taxon, fill = taxon)) +
  geom_point()+
  geom_smooth(method = "lm")+
  scale_color_brewer(palette = "Paired")+
  scale_fill_brewer(palette = "Paired")+
  ylab("simulated proportion of woody cover")+
  xlab(expression("simulated "*delta^{13}*"C"))+
  theme_classic(base_size = 16)
ggsave("figures/simulated_taxa_wc_vs_d13C.png", width = 9, height = 7)
```

## 4. Use a varying effects model to examine how well d13C can predict woody cover.

### Model syntax:

Woody cover is modeled as a Beta dstribution, with mean ${\mu}$ and variance ${\phi}$.

$$ WC \sim \mathrm{Beta}({\mu_i}, {\phi}) $$
The estimated ${\mu}$ based on each individual's ${\delta}^{13}C$ value is modeled using the logit regression below. The intercept (${\alpha}$) and slope (${\beta}$) terms vary by taxon. 

$$ \mathrm{logit}(\mu_i) = \alpha_{[taxon]_i} + \beta_{[taxon]_i}*{\delta}^{13}C_i $$
The varying slope and intercept terms are distributed according to a multivariate normal distribution, centered at the mean intercept and slope, and varying according to the ovariance matrix, ${\Sigma}$.

$$ \begin{bmatrix} \alpha_{[taxon]_i}\\ \beta_{[taxon]_i} \end{bmatrix} \sim \mathrm{MVNormal}(\begin{bmatrix} \alpha\\ \beta \end{bmatrix}, \mathbf{\Sigma})$$

That covariance matrix can be factored out into the two individual covariance matrices for the slope and intercept parameters, where the variance for each term is represented by its standard deviation (${\sigma}$), multiplied by the correlation matrix, R. 

$$ \mathbf{\Sigma} = \begin{bmatrix} \sigma_{\alpha} & 0\\ 0 & \sigma_{\alpha} \end{bmatrix} \mathbf{R} \begin{bmatrix} \sigma_{\beta} & 0\\ 0 & \sigma_{\beta} \end{bmatrix}  $$

The priors for each of these parameters are as follows:
$$ \alpha \sim \mathrm{Normal}(0, 0.5) $$
$$ \beta \sim Normal(0, 0.5)$$
$$  \sigma_{\alpha} \sim Gamma(2, 10)$$
$$  \sigma_{\beta} \sim Gamma(2, 10)$$
$$ \phi \sim Gamma(15, 2)$$
$$ \mathbf{R} \sim \mathrm{LKJcorr}(2) $$

### Run the simulated data model

Set the model syntax and priors:
```{r}
# model formula 
iso.sim <- bf(wc_sim ~ 1 + d13C_scaled + (1 + d13C_scaled | taxon)) 

# default priors
get_prior(formula = iso.sim, data = dsim, family = Beta())

# set the priors
iso.sim.priors <- c(prior("normal(0, 0.5)", class = "b"), prior("gamma(15, 2)", class = "phi"), prior("normal(0, 0.5)", class = "Intercept"), prior("gamma(2,10)", class = "sd"), prior("lkj(2)", class = "cor"))
```

Run a prior predictive check:
```{r}
# prior predictive check
iso.sim.pp.check <- brm(
   iso.sim,
   prior <- iso.sim.priors,
   sample_prior = 'only',
   data = dsim,
   family = Beta(),
   iter = 1000,
   chains = 4, 
   cores = 4,
   file = "models/simulations/iso_sim_pp_check")

iso.sim.pp.check
plot(iso.sim.pp.check, nvariables=3, ask=F)

set.seed(2474)
pp_check(iso.sim.pp.check, ndraws = 100)
```

Run the model on the data:
```{r}
iso.sim.model <- brm(
   iso.sim,
   prior <- iso.sim.priors,
   data = dsim,
   family = Beta(),
   chains = 4, 
   cores = 4,
   iter = 3000,
   control = list(adapt_delta = 0.99),
   file = "models/simulations/iso_sim")

iso.sim.model
```

Examine the output. Is the model estimating the parameters as expected based on the data generation process?
```{r}
iso.sim.model$fit
```

Posterior predictive check:
```{r}
pp_check(iso.sim.model, ndraws = 300)
```

### Visualize model output

Fixed effects plot:
```{r, warning=FALSE, message=FALSE}
iso.sim.model.post <- as_draws_df(iso.sim.model) # sample 2000 posterior draws
iso.sim.model.post %>% 
  dplyr::select(b_Intercept:cor_taxon__Intercept__d13C_scaled) %>%
  pivot_longer(b_Intercept:cor_taxon__Intercept__d13C_scaled) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for fixed effects") +
  theme_classic(base_size = 12)
ggsave("figures/fixed_effects_varying_effects_simulated_data.png", width = 9, height = 7)
```

Varying intercepts:
```{r, warning=FALSE, message=FALSE}
iso.sim.model.post %>% 
  dplyr::select(`r_taxon[A,Intercept]`:`r_taxon[J,Intercept]`) %>%
  pivot_longer(`r_taxon[A,Intercept]`:`r_taxon[J,Intercept]`) %>%
  mutate(name = substr(name, 9, nchar(name)-11)) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab("taxon") +
  xlab("coefficients for varying intercepts (log-odds)") +
  theme_classic(base_size = 12)
ggsave("figures/varying_intercept_coefficients_varying_slopes_simulated_data.png", width = 8, height = 7)
```

Varying slopes:
```{r, warning=FALSE, message=FALSE}
iso.sim.model.post %>% 
  dplyr::select(`r_taxon[A,d13C_scaled]`:`r_taxon[J,d13C_scaled]`) %>%
  pivot_longer(`r_taxon[A,d13C_scaled]`:`r_taxon[J,d13C_scaled]`) %>%
  mutate(name = substr(name, 9, nchar(name)-13)) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab("taxon") +
  xlab("coefficients for varying slopes (log-odds)") +
  theme_classic(base_size = 12)
ggsave("figures/slope_coefficients_varying_slopes_simulated_data.png", width = 8, height = 7)
```

How well does the model predict?
```{r, results='hide', message=FALSE}
iso.sim.model.post.pred <- posterior_epred(iso.sim.model) # generate posterior predictions for each site
iso.sim.model.post.pred<- t(iso.sim.model.post.pred)
iso.sim.model.post.pred <- bind_cols(dsim[,c("site", "taxon")], iso.sim.model.post.pred)
iso.sim.model.post.pred <- pivot_longer(iso.sim.model.post.pred, cols = c(3:6002))
```

```{r, fig.height=9, fig.width=7, echo=FALSE}
full_predictions <- iso.sim.model.post.pred %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = "")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_blank(), legend.position = "bottom")
ggsave(full_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data.png", height = 9, width = 7)
```

Coloring by taxon:
```{r, fig.height=9, fig.width=8, echo=FALSE}
full_predictions_by_taxon <- iso.sim.model.post.pred %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_brewer(palette = "Paired", name = "estimated value\nby taxon") +
  scale_color_brewer(palette = "Paired", name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_text(size = 12))
ggsave(full_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_by_taxon.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(full_predictions | full_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_compare.pdf", height = 9, width = 14)
```

Plot "very strong" slopes only
```{r, fig.height=9, fig.width=7, echo=FALSE}
strong_predictions_by_taxon <- iso.sim.model.post.pred %>% filter(taxon %in% c("A", "B", "C")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_brewer(palette = "Paired", name = "estimated value\nby taxon") +
  scale_color_brewer(palette = "Paired", name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_text(size = 12))
ggsave(strong_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_strong_slopes_by_taxon.png", height = 9, width = 7)

strong_predictions <- iso.sim.model.post.pred %>% filter(taxon %in% c("A", "B", "C")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_blank())
ggsave(strong_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data_strong_slopes.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(strong_predictions | strong_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_strong_slopes_compare.pdf", height = 9, width = 14)
```

Plot "mild" slopes only
```{r, fig.height=9, fig.width=7, echo=FALSE}
mild_predictions_by_taxon <- iso.sim.model.post.pred %>% filter(taxon %in% c("D", "E", "F")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_manual(values = c("#33a02c","#fb9a99","#e31a1c"), name = "estimated value\nby taxon") +
  scale_color_manual(values = c("#33a02c","#fb9a99","#e31a1c"), name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_text(size = 12))
ggsave(mild_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_mild_slopes_by_taxon.png", height = 9, width = 7)

mild_predictions <- iso.sim.model.post.pred %>% filter(taxon %in% c("D", "E", "F")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_blank(), legend.position = "bottom")
ggsave(mild_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data_mild_slopes.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(mild_predictions | mild_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_mild_slopes_compare.pdf", height = 9, width = 14)
```

Plot "weak" slopes only
```{r, fig.height=9, fig.width=7, echo=FALSE}
weak_predictions_by_taxon <- iso.sim.model.post.pred %>% filter(taxon %in% c("G", "H", "I", "J")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_manual(values = c("#fdbf6f","#ff7f00","#cab2d6","#6a3d9a"), name = "estimated value\nby taxon") +
  scale_color_manual(values = c("#fdbf6f","#ff7f00","#cab2d6","#6a3d9a"), name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_text(size = 12), legend.position = "bottom")
ggsave(weak_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_weak_slopes_by_taxon.png", height = 9, width = 7)

weak_predictions <- iso.sim.model.post.pred %>% filter(taxon %in% c("G", "H", "I", "J")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_blank(), legend.position = "bottom")
ggsave(weak_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data_weak_slopes.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(weak_predictions | weak_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_weak_slopes_compare.pdf", height = 9, width = 14)
```

# Model with real data

Based on the model using simulated data, and previous work examining the relationship between d13C and woody cover, we expect that the accuracy of predictions of woody cover will depend on the community composition and which taxa are sampled. 

```{r}
# create new column in dataset for taxonomy used in this analysis
iso_data$taxon_model <- NA
iso_data[iso_data$Family == "Bovidae",]$taxon_model <- iso_data[iso_data$Family == "Bovidae",]$Tribe
iso_data[iso_data$Family != "Bovidae",]$taxon_model <- iso_data[iso_data$Family != "Bovidae",]$Family
```

```{r}
# make pretreatment_detail column into a factor 
iso_data$Pretreatment_Detail <- as.factor(iso_data$Pretreatment_Detail)
levels(iso_data$Pretreatment_Detail) # 7 categories in the detailed method column
```

## Specify the model
```{r}
# model formula - requires a non-linear parameterization because of the two types of varying effects
iso.model.formula <- bf(pct_wc_mean ~ 0 + d + p,
                          d ~ 1 + d13C_scaled + (1 + d13C_scaled | taxon_model),
                          p ~ 0 + Pretreatment_Detail,
                          nl = TRUE)

# default priors
get_prior(formula = iso.model.formula, data = iso_data, family = Beta())

# set the priors 
iso.model.priors <- c(prior("normal(0, 0.5)", class = "b", nlpar = "d"), prior("gamma(15, 2)", class = "phi"), prior("gamma(2,10)", class = "sd", nlpar = "d"), prior("lkj(2)", class = "cor"), prior("normal(0, 0.5)", class = "b", nlpar = "p"))
```

## Prior predictive check
```{r}
# prior predictive check
iso.model.pp.check <- brm(
   iso.model.formula,
   prior <- iso.model.priors,
   sample_prior = 'only',
   data = iso_data,
   family = Beta(),
   iter = 1000,
   chains = 4, 
   cores = 4,
   file = "models/with_data/iso_model_pp_check")

iso.model.pp.check
plot(iso.model.pp.check, nvariables=3, ask=F)

set.seed(23674)
pp_check(iso.model.pp.check , ndraws = 100)
ggsave("figures/prior_predictive_check_iso_model.png", bg = "white")
```

## Sample posterior
```{r}
iso.model <- brm(
   iso.model.formula,
   prior <- iso.model.priors,
   data = iso_data,
   family = Beta(),
   chains = 4, 
   cores = 4,
   iter = 3000,
   control = list(adapt_delta = 0.99),
   file = "models/with_data/iso_model")

iso.model
```

```{r}
iso.model$fit
```

Posterior predictive check:
```{r}
pp_check(iso.model, ndraws = 300)
ggsave("figures/posterior_predictive_check_iso_model.png", bg = "white")
```

## Visualize model output

Fixed effects:
```{r, warning=FALSE, message=FALSE}
iso.model.post <- as_draws_df(iso.model) # sample 2000 posterior draws

# fixed effects minus pretreatment coefficients
iso.model.post %>% 
  dplyr::select(b_d_Intercept, b_d_d13C_scaled, sd_taxon_model__d_Intercept, sd_taxon_model__d_d13C_scaled, cor_taxon_model__d_Intercept__d_d13C_scaled) %>%
  pivot_longer(cols = c(b_d_Intercept, b_d_d13C_scaled, sd_taxon_model__d_Intercept, sd_taxon_model__d_d13C_scaled, cor_taxon_model__d_Intercept__d_d13C_scaled)) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for fixed effects (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/fixed_effects_iso_model.png", width = 9, height = 7)

# pretreatment effects
iso.model.post %>% 
  dplyr::select(b_p_Pretreatment_Detail0.1M_buff_AA:b_p_Pretreatment_DetailNone) %>%
  pivot_longer(b_p_Pretreatment_Detail0.1M_buff_AA:b_p_Pretreatment_DetailNone) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for pretreatment effects (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/pretreatment_effects_iso_model.png", width = 9, height = 7)
```

Varying intercepts:
```{r, warning=FALSE, message=FALSE}
iso.model.post %>% 
  dplyr::select(`r_taxon_model__d[Aepycerotini,Intercept]`:`r_taxon_model__d[Viverridae,Intercept]`) %>%
  pivot_longer(`r_taxon_model__d[Aepycerotini,Intercept]`:`r_taxon_model__d[Viverridae,Intercept]`) %>%
  mutate(name = substr(name, 18, nchar(name)-11)) %>%
  mutate(name = fct_reorder(name, desc(value))) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for varying intercepts (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/varying_intercept_coefficients_iso_model.png", width = 9, height = 7)
```

Varying slopes:
```{r, warning=FALSE, message=FALSE}
iso.model.post %>% 
  dplyr::select(`r_taxon_model__d[Aepycerotini,d13C_scaled]`:`r_taxon_model__d[Viverridae,d13C_scaled]`) %>%
  pivot_longer(`r_taxon_model__d[Aepycerotini,d13C_scaled]`:`r_taxon_model__d[Viverridae,d13C_scaled]`) %>%
  mutate(name = substr(name, 18, nchar(name)-13)) %>%
  mutate(name = fct_reorder(name, desc(value))) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for varying slopes (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/varying_slopes_iso_model.png", width = 9, height = 7)
```

Extract posterior predictions of woody cover:
```{r, results='hide', message=FALSE}
iso.model.post.pred <- posterior_epred(iso.model) # generate posterior predictions for each specimen
iso.model.post.pred<- t(iso.model.post.pred)
iso.model.post.pred <- bind_cols(iso_data[,c("locality", "Family", "taxon_model", "Sample_ID", "COUNTRY", "pct_wc_mean")], iso.model.post.pred)
iso.model.post.pred <- pivot_longer(iso.model.post.pred, cols = c(7:6006))
```

Plot overall predictions for each country:
```{r, echo=FALSE}
# uses custom "country_plot" function

post.cameroon <- country_plot(posterior_preds = iso.model.post.pred, country = "Cameroon", real_data = iso_data)
ggsave("figures/post_cameroon.png", post.cameroon, width = 7, height = 4)

post.drc <- country_plot(posterior_preds = iso.model.post.pred, country = "DRC", real_data = iso_data)
ggsave("figures/post_drc.png", post.drc, width = 8, height = 6)

post.ethiopia <- country_plot(posterior_preds = iso.model.post.pred, country = "Ethiopia", real_data = iso_data)
ggsave("figures/post_ethiopia.png", post.ethiopia, width = 9, height = 8)

post.gabon <- country_plot(posterior_preds = iso.model.post.pred, country = "Gabon", real_data = iso_data)
post.gabon <- post.gabon + scale_y_discrete(labels = c("La Lopé")) # fix special character in site name
ggsave("figures/post_gabon.png", post.gabon, width = 7, height = 4)

post.kenya <- country_plot(posterior_preds = iso.model.post.pred, country = "Kenya", real_data = iso_data)
ggsave("figures/post_kenya.png", post.kenya, width = 10, height = 10)

post.liberia <- country_plot(posterior_preds = iso.model.post.pred, country = "Liberia", real_data = iso_data)
ggsave("figures/post_liberia.png", post.liberia, width = 7, height = 4)

post.moz <- country_plot(posterior_preds = iso.model.post.pred, country = "Mozambique", real_data = iso_data)
ggsave("figures/post_mozambique.png", post.moz, width = 7, height = 4)

post.sa <- country_plot(posterior_preds = iso.model.post.pred, country = "South Africa", real_data = iso_data)
ggsave("figures/post_southafrica.png", post.sa, width = 9, height = 8)

post.tanzania <- country_plot(posterior_preds = iso.model.post.pred, country = "Tanzania", real_data = iso_data)
ggsave("figures/post_tanzania.png", post.tanzania, width = 8, height = 6)

post.uganda <- country_plot(posterior_preds = iso.model.post.pred, country = "Uganda", real_data = iso_data)
ggsave("figures/post_uganda.png", post.uganda, width = 9, height = 6)
```

Eastern African localities:
```{r}
((post.ethiopia / post.tanzania / post.uganda) + plot_layout(axis_titles = "collect", heights = c(2,1,1)) | post.kenya ) + plot_layout(guides = "collect")
ggsave("figures/post_predictions_eastern_africa.png", height = 12, width = 16)
```

Western, central, and southern African localities:
```{r}
((post.cameroon / post.gabon / post.liberia / post.drc / post.moz) + plot_layout(axis_titles = "collect", heights = c(1,1,1,3,1)) | post.sa) + plot_layout(guides = "collect")
ggsave("figures/post_predictions_western_central_southern_africa.png", height = 11, width = 14)
```

#### Plot by taxon
```{r, echo=F}
post.cameroon.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Cameroon", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_cameroon.png", post.cameroon.taxa, height = 4, width = 7)

post.drc.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "DRC", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_drc.png", post.drc.taxa, height = 6, width = 8)

post.ethio.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Ethiopia", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_ethiopia.png", post.ethio.taxa, height = 8, width = 10)

post.gabon.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Gabon", real_data = iso_data)
post.gabon.taxa <- post.gabon.taxa + scale_y_discrete(labels = c("La Lopé")) # fix special character in site name
ggsave("figures/posterior_predictions_iso_model_by_taxon_gabon.png", post.gabon.taxa, height = 4, width = 7)

post.kenya.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Kenya", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_kenya.png", post.kenya.taxa, height = 12, width = 13)

post.liberia.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Liberia", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_liberia.png", post.liberia.taxa, height = 4, width = 7)

post.moz.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Mozambique", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_mozambique.png", post.moz.taxa, height = 4, width = 7)

post.sa.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "South Africa", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_sa.png", post.sa.taxa, height = 8, width = 9)

post.tanzania.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Tanzania", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_tanzania.png", post.tanzania.taxa, height = 6, width = 8)

post.uganda.taxa <- country_taxon_plot(posterior_preds = iso.model.post.pred, country = "Uganda", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_uganda.png", post.uganda.taxa, width = 9, height = 6)
```


#### Plot bovids by tribe

```{r}
# can't plot Cameroon or Liberia because there are no bovids in either sample

post.drc.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "DRC", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_drc.png", post.drc.bovids, height = 6, width = 8)

post.ethio.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "Ethiopia", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_ethiopia.png", post.ethio.bovids, height = 8, width = 10)

post.gabon.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "Gabon", real_data = iso_data)
post.gabon.bovids <- post.gabon.bovids + scale_y_discrete(labels = c("La Lopé")) # fix special character
ggsave("figures/posterior_predictions_iso_model_bovids_gabon.png", post.gabon.bovids, height = 6, width = 8)

post.kenya.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "Kenya", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_kenya.png", post.kenya.bovids, height = 12, width = 13)

post.moz.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "Mozambique", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_mozambique.png", post.moz.bovids, height = 4, width = 8)

post.sa.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "South Africa", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_sa.png", post.sa.bovids, height = 8, width = 10)

post.tanzania.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "Tanzania", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_tanzania.png", post.tanzania.bovids, height = 5, width = 8)

post.uganda.bovids <- country_bovid_plot(posterior_preds = iso.model.post.pred, country = "Uganda", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_uganda.png", post.uganda.bovids, width = 9, height = 6)
```

### Summary plot of how predictions at all sites compare to real mean woody cover

```{r}
# summary statistics for posterior predictions by site
summary.iso.model <- iso.model.post.pred %>% group_by(locality) %>% mutate(mean_wc_pred = mean(value), med_wc_pred = median(value), var_wc_pred = var(value), sd_wc_pred = sd(value)) %>% dplyr::select(locality, mean_wc_pred, med_wc_pred, var_wc_pred, sd_wc_pred) %>% unique()

# join summary statistics with data from woody cover dataset
summary.iso.model <- left_join(summary.iso.model, iso_data[,c("locality", "wc_mean", "wc_median", "wc_variance", "wc_stdev", "AREA..sq.km.")], by = "locality") %>% unique()

# convert woody cover mean and median values to proportion
summary.iso.model[,c("wc_mean", "wc_median")] <- summary.iso.model[,c("wc_mean", "wc_median")]/100
```

```{r}
ggplot(data = summary.iso.model, aes(y = mean_wc_pred, x = wc_mean, fill = wc_mean)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2)+
  geom_errorbar(aes(ymin=mean_wc_pred-sd_wc_pred, ymax = mean_wc_pred+sd_wc_pred))+
  geom_point(shape = 21, size = 4)+
  scale_fill_distiller(palette = "BrBG", direction = 1, name = NULL)+
  xlim(0,1)+
  ylim(0,1)+
  labs(y = "posterior predicted mean woody cover", x = "mean woody cover")+
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom")
ggsave("figures/posterior_vs_actual_woody_cover_model.png", height = 8, width = 9)
```
