---
title: "Predicting woody cover with enamel carbon isotopes using Bayesian multi-level modeling"
author: "Maddie Kelly"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages & custom functions
```{r}
library(tidyverse)
library(ggmap)
library(brms)
library(tidybayes)
library(bayesplot)
library(patchwork)
library(ggnewscale)
library(ggstar)
```

```{r}
# custom functions for plotting 
source(file = "code/custom_functions.R", local = T)
```

# Import dataset
```{r}
iso_data <- read_csv("data/modern_enamel_d13C.csv")

# the d13C_corrected column contains the d13C values corrected to a pre-industrial atmospheric d13C, which are used for all analyses

# center and scale d13C values 
iso_data$d13C_scaled <- scale(iso_data$d13C_corrected)
```

## Figure 1A: Map of sites 
```{r mapping}
# use stadia maps
register_stadiamaps(key = "YOUR-KEY-HERE") # input personal Stadia maps API key to generate map 
terrainmap <- get_map(location = c(left = -20, bottom = -37, right = 50, top = 23), zoom = 4, maptype = "stamen_terrain_background", source = "stadia")
finelines <- get_map(location = c(left = -20, bottom = -37, right = 50, top = 23), zoom = 4, maptype = "stamen_toner_lines", source = "stadia")

terrainmap_attributes <- attributes(terrainmap)
terrainmap_transparent <- matrix(adjustcolor(terrainmap, 
                                                alpha.f = 0.9), 
                                    nrow = nrow(terrainmap))
attributes(terrainmap_transparent) <- terrainmap_attributes

# plotting all of the localities included in the analysis 
isotope_data_map <- ggmap(terrainmap_transparent)+
  inset_ggmap(finelines)+ # sticks the country lines on top 
  geom_point(data = iso_data, aes(x = Longitude_updated, y = Latitude_updated), fill = "#A5506DFF", size = 3, shape = 24)+
  scale_y_continuous(position = "right", expand = 0)+
  labs(x = "Longitude", y = "Latitude", title = "Modern enamel carbon isotope data localities")+
  annotate("text",label = "© Stadia Maps © Stamen Design\n© OpenMapTiles © OpenStreetMap contributors", x = -18, y = -33, size = 3, hjust=0)+
  theme(axis.text = element_text(size = 10), plot.title = element_text(hjust = 0.5, size = 14))

isotope_data_map
ggsave(isotope_data_map, filename = "figures/isotope_data_map.pdf")

```

## Fiure 1B: Barplot of sample size by family:
```{r}
iso_data %>% group_by(Family) %>% tally() %>% 
  ggplot() +
  geom_col(aes(x = reorder(Family, n, decreasing = T), y = n, color = Family, fill = Family), alpha = 0.4, linewidth = 2) +
  family_colors()+
  family_fills()+
  labs(x = "Family", y = "Number of samples")+
  theme_classic(base_size = 18)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/d13C_family_plot.png", width = 9, height = 7)
```

## (Supplementary Figure 1) Plot distributions of woody cover at each site 

```{r}
wc_data_gis <- read.csv("data/sites_in_analysis_histo.csv")
wc_data_gis <- wc_data_gis %>% select(name, parent_iso, HISTO_1:HISTO_99)
wc_data_gis <- wc_data_gis %>% pivot_longer(cols = c(HISTO_1:HISTO_99), names_to = "wc_value")
wc_data_gis <- wc_data_gis %>% filter(value != 0)
wc_data_gis <- wc_data_gis %>% uncount(value)
wc_data_gis$wc_value <- sub("HISTO_", "", wc_data_gis$wc_value)
wc_data_gis$wc_value <- as.numeric(wc_data_gis$wc_value)

real_cameroon <- wc_data_gis %>% filter(parent_iso == "CAM") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Cameroon")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_drc <- wc_data_gis %>% filter(parent_iso == "COD") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("DRC")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_ethiopia <- wc_data_gis %>% filter(parent_iso == "ETH") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Ethiopia")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_gabon <- wc_data_gis %>% filter(parent_iso == "GAB") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    scale_y_discrete(labels = c("La Lopé"))+
    ylab(NULL)+
    ggtitle("Gabon")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_kenya <- wc_data_gis %>% filter(parent_iso == "KEN") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Kenya")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_liberia <- wc_data_gis %>% filter(parent_iso == "LIB") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Liberia")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_moz <- wc_data_gis %>% filter(parent_iso == "MOZ") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Mozambique")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_sa <- wc_data_gis %>% filter(parent_iso == "ZAF") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("South Africa")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_tanzania <- wc_data_gis %>% filter(parent_iso == "TZA") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Tanzania")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())

real_uganda <- wc_data_gis %>% filter(parent_iso == "UGA") %>%
    ggplot()+
    stat_halfeye(aes(x = wc_value/100, y = name), color = "darkgreen", fill = alpha("darkgreen", alpha = 0.6), point_size = 3, point_interval = "median_qi", .width = 0.89, normalize = "groups")+
    xlab("proportion woody cover") +
    xlim(0,1)+
    ylab(NULL)+
    ggtitle("Uganda")+
    theme_classic(base_size = 16)+
    theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5), legend.title = element_blank())
```


```{r, fig.height=20, fig.width=16}
((real_ethiopia / real_tanzania / real_uganda / real_drc / real_moz / real_cameroon / real_gabon / real_liberia) + plot_layout(axis_titles = "collect", heights = c(7,3,4,4,1,1,1,1))) | ((real_kenya / real_sa) + plot_layout(axis_titles = "collect", heights = c(21,9)))
ggsave("figures/real_woody_cover_distributions.pdf", width = 16, height = 20)
```

## (Supplementary Figure 2) Plot mean versus median woody cover at each site 
```{r}
ggplot(data = iso_data, aes(x = wc_mean, y = wc_median))+
  geom_point(size = 2)+
  geom_abline(aes(slope = 1, intercept = 0))+
  labs(x = "Mean percent woody cover", y = "Median percent woody cover")+
  theme_classic(base_size = 16)
ggsave("figures/mean_vs_median_wc_data.png", width = 7, height = 7)
```

Summary statistics for median woody cover across all sites:
```{r}
site_wc <- iso_data %>% select(locality, wc_median) %>% distinct()
summary(site_wc$wc_median)
```

# Model with simulated data

## 1. Simulate slope and intercept parameters for a group of 10, random hypothetical taxa. 

Three taxa will be given very strong slopes, three given mild slopes, and the remaining four given weak slopes, to simulate the variation in the relationship between d13C and woody cover in real mammal taxa. The intercepts will be drawn at random from a Gaussian distribution with a mean 0.5 and standard deviation 0.25. In this case, the slopes and intercepts are not explicitly modeled to covary. 
```{r}
ggplot()+
  geom_function(fun = dnorm, args = list(mean = -0.05, sd = 0.02), aes(color = "very strong"), linewidth = 2)+
  geom_function(fun = dnorm, args = list(mean = -0.2, sd = 0.05), aes(color = "mild"), linewidth = 2)+
  geom_function(fun = dnorm, args = list(mean = -0.5, sd = 0.1), aes(color = "weak"), linewidth = 2)+
  scale_color_manual(values = c("very strong" = "mediumpurple", "mild" = "coral2", "weak" = "darkgreen"), name = "varying slopes")+
  xlim(-1,0.5)+
  theme_classic()
ggsave("figures/simulation_slope_distributions.png")

set.seed(2840) # set seed for reproducibility
sim_taxa <- data.frame(taxon = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"))
sim_taxa <- sim_taxa %>% mutate(intercept = rnorm(n(), mean = 0.5, sd = 0.25)) # generate random intercept values for each taxon
sim_taxa$slope <- NA
sim_taxa[1:3,] <- sim_taxa[1:3,] %>% 
  mutate(slope = rnorm(n(), mean = -0.05, sd = 0.02)) # generate a set of very strong slopes
sim_taxa[4:6,] <- sim_taxa[4:6,] %>%
  mutate(slope = rnorm(n(), mean = -0.2, sd = 0.05)) # generate a set of weaker slopes
sim_taxa[7:10,] <- sim_taxa[7:10,] %>%
  mutate(slope = rnorm(n(), mean = -0.5, sd = 0.1)) # generate a set of very weak slopes
```

## 2. Simulate a set of sites (n = 20) with random woody cover values, and with 10 of each taxon at each site (n = 100 individuals per site). 

Using the taxon specific slope and intercept parameters, paired with the site's woody cover value, generate simulated d13C values for each individual at each site. 
```{r}
set.seed(2840) # for reproducibility
nsites = 20
mu = 0.35
phi = 7.5
dsim <- tibble(site = 1:nsites,
               wc_sim = rbeta(n = nsites, shape1 = mu*phi, shape2 = (1-mu)*phi)) # draw woody cover values from this Beta distribution; mu and phi (mean and variance) of the beta distribution are converted to shape1 and shape2 parameters using the equations shape1 = mu * phi and shape2 = (1-mu)*phi.

# repeat rows of simulated woody cover dataset to generate multiple observations for each "site". 
dsim <- dsim %>% slice(rep(1:n(), each = 10))
dsim$site <- as.factor(dsim$site)

# simulate "carbon isotope" values for each hypothetical taxon, for each site, using the woody cover value and intercept and beta parameters set for each taxon
for (i in 1:nrow(sim_taxa)) {
  for (j in 1:nrow(dsim)) { 
    dsim[j,paste("d13C_sim_", sim_taxa[i,"taxon"], sep = "")] <- rnorm(1, mean = (dsim[j,]$wc_sim - sim_taxa[i,"intercept"]) / sim_taxa[i, "slope"], sd = 1.5)
  }
}

# does that generate the relationships I want to see?
plot(density(dsim$d13C_sim_A))
plot(dsim$wc_sim~dsim$d13C_sim_A)

# pivot the data so there's only one d13C column and another column indicating taxon
dsim <- dsim %>% pivot_longer(d13C_sim_A:d13C_sim_J)

# rename columns and column values for taxa
colnames(dsim)[colnames(dsim) %in% c("name", "value")] <- c("taxon", "d13C_sim")
dsim$taxon <- sub("d13C_sim_", "", dsim$taxon)

# scale d13C values 
dsim$d13C_scaled <- scale(dsim$d13C_sim)

```

## 3. (Supplementary Figure 3) Visualize these simulated relationships to make sure that the data generation process generated the expected results.
```{r}
ggplot(data = dsim, aes(x = d13C_sim, y = wc_sim, color = taxon, fill = taxon)) +
  geom_point()+
  geom_smooth(method = "lm")+
  scale_color_brewer(palette = "Paired")+
  scale_fill_brewer(palette = "Paired")+
  ylab("simulated proportion of woody cover")+
  xlab(expression("simulated "*delta^{13}*"C"))+
  theme_classic(base_size = 16)
ggsave("figures/simulated_taxa_wc_vs_d13C.png", width = 9, height = 7)
```

## 4. Use a varying effects model to examine how well d13C can predict woody cover.

### Model syntax:

Woody cover is modeled as a Beta dstribution, with mean ${\mu}$ and variance ${\phi}$.

$$ WC \sim \mathrm{Beta}({\mu_i}, {\phi}) $$
The estimated ${\mu}$ based on each individual's ${\delta}^{13}C$ value is modeled using the logit regression below. The intercept (${\alpha}$) and slope (${\beta}$) terms vary by taxon. 

$$ \mathrm{logit}(\mu_i) = \alpha_{[taxon]_i} + \beta_{[taxon]_i}*{\delta}^{13}C_i $$
The varying slope and intercept terms are distributed according to a multivariate normal distribution, centered at the mean intercept and slope, and varying according to the covariance matrix, ${\Sigma}$.

$$ \begin{bmatrix} \alpha_{[taxon]_i}\\ \beta_{[taxon]_i} \end{bmatrix} \sim \mathrm{MVNormal}(\begin{bmatrix} \alpha\\ \beta \end{bmatrix}, \mathbf{\Sigma})$$

That covariance matrix can be factored out into the two individual covariance matrices for the slope and intercept parameters, where the variance for each term is represented by its standard deviation (${\sigma}$), multiplied by the correlation matrix, R. 

$$ \mathbf{\Sigma} = \begin{bmatrix} \sigma_{\alpha} & 0\\ 0 & \sigma_{\alpha} \end{bmatrix} \mathbf{R} \begin{bmatrix} \sigma_{\beta} & 0\\ 0 & \sigma_{\beta} \end{bmatrix}  $$

The priors for each of these parameters are as follows:
$$ \alpha \sim \mathrm{Normal}(0, 0.5) $$
$$ \beta \sim Normal(0, 0.5)$$
$$  \sigma_{\alpha} \sim Gamma(2, 10)$$
$$  \sigma_{\beta} \sim Gamma(2, 10)$$
$$ \phi \sim Gamma(15, 2)$$
$$ \mathbf{R} \sim \mathrm{LKJcorr}(2) $$

### Run the simulated data model

Set the model syntax and priors:
```{r}
# model formula 
iso.sim <- bf(wc_sim ~ 1 + d13C_scaled + (1 + d13C_scaled | taxon)) 

# default priors
get_prior(formula = iso.sim, data = dsim, family = Beta())

# set the priors
iso.sim.priors <- c(prior("normal(0, 0.5)", class = "b"), prior("gamma(15, 2)", class = "phi"), prior("normal(0, 0.5)", class = "Intercept"), prior("gamma(2,10)", class = "sd"), prior("lkj(2)", class = "cor"))
```

Run a prior predictive check:
```{r}
# prior predictive check
iso.sim.pp.check <- brm(
   iso.sim,
   prior <- iso.sim.priors,
   sample_prior = 'only',
   data = dsim,
   family = Beta(),
   iter = 1000,
   chains = 4, 
   cores = 4,
   file = "models/simulations/iso_sim_pp_check")

iso.sim.pp.check
plot(iso.sim.pp.check, nvariables=3, ask=F)

set.seed(2474)
pp_check(iso.sim.pp.check, ndraws = 100)
```

Run the model on the simulated data:
```{r}
iso.sim.model <- brm(
   iso.sim,
   prior <- iso.sim.priors,
   data = dsim,
   family = Beta(),
   chains = 4, 
   cores = 4,
   iter = 3000,
   control = list(adapt_delta = 0.99),
   file = "models/simulations/iso_sim")

iso.sim.model
```
 
Examine the output. Is the model estimating the parameters as expected based on the data generation process?
```{r}
iso.sim.model$fit
```

Summed intercept and slope parameters for simulated taxa:
```{r}
varef_sim_model <- coef(iso.sim.model, summary = F)$taxon
```

Posterior predictive check:
```{r}
pp_check(iso.sim.model, ndraws = 300)
```

### Visualize model output

Fixed effects plot: 
```{r, warning=FALSE, message=FALSE}
iso.sim.model.post <- as_draws_df(iso.sim.model) # sample 2000 posterior draws
iso.sim.model.post %>% 
  dplyr::select(b_Intercept:cor_taxon__Intercept__d13C_scaled) %>%
  pivot_longer(b_Intercept:cor_taxon__Intercept__d13C_scaled) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for fixed effects") +
  theme_classic(base_size = 12)
ggsave("figures/fixed_effects_varying_effects_simulated_data.png", width = 9, height = 7)
```

#### Figure 2A: Varying intercepts
```{r, warning=FALSE, message=FALSE}
# summed with fixed intercept
varef_sim_model[,,"Intercept"] %>%
   as_data_frame() %>%
   gather() %>%
   ggplot() +
   stat_pointinterval(aes(x = value, y = key), color = "#300948")+
   geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
   ylab(NULL) +
   xlab("coefficients for varying intercepts (log-odds)") +
   theme_classic(base_size = 16)
ggsave("figures/SUM_varying_intercept_coefficients_varying_slopes_simulated_data.png", width = 8, height = 7)
```

#### Figure 2B: Varying slopes:
```{r, warning=FALSE, message=FALSE}
# summed with fixed slope
varef_sim_model[,,"d13C_scaled"] %>%
   as_data_frame() %>%
   gather() %>%
   ggplot() +
   stat_pointinterval(aes(x = value, y = key), color = "#300948")+
   geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
   ylab(NULL) +
   xlab("coefficients for varying slopes (log-odds)") +
   theme_classic(base_size = 16)
ggsave("figures/SUM_slope_coefficients_varying_slopes_simulated_data.png", width = 8, height = 7)
```

How well does the model predict?
```{r, results='hide', message=FALSE}
iso.sim.model.post.pred <- posterior_epred(iso.sim.model) # generate posterior predictions for each site
iso.sim.model.post.pred<- t(iso.sim.model.post.pred)
iso.sim.model.post.pred <- bind_cols(dsim[,c("site", "taxon")], iso.sim.model.post.pred)
iso.sim.model.post.pred <- pivot_longer(iso.sim.model.post.pred, cols = c(3:6002))
```

#### Figure 3A: Posterior predictions
```{r, fig.height=9, fig.width=7, echo=FALSE}
full_predictions <- iso.sim.model.post.pred %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = "")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_blank(), legend.position = "bottom")
ggsave(full_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data.png", height = 9, width = 7)
```

#### Figure 3B: Posterior predictions, by taxon
```{r, fig.height=9, fig.width=8, echo=FALSE}
full_predictions_by_taxon <- iso.sim.model.post.pred %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_brewer(palette = "Paired", name = "estimated value\nby taxon") +
  scale_color_brewer(palette = "Paired", name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_text(size = 12))
ggsave(full_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_by_taxon.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(full_predictions | full_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_compare.pdf", height = 9, width = 14)
```

#### Figure 3C-D: Plot "very strong" slopes only
```{r, fig.height=9, fig.width=7, echo=FALSE}
strong_predictions_by_taxon <- iso.sim.model.post.pred %>% filter(taxon %in% c("A", "B", "C")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_brewer(palette = "Paired", name = "estimated value\nby taxon") +
  scale_color_brewer(palette = "Paired", name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_text(size = 12))
ggsave(strong_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_strong_slopes_by_taxon.png", height = 9, width = 7)

strong_predictions <- iso.sim.model.post.pred %>% filter(taxon %in% c("A", "B", "C")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_blank())
ggsave(strong_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data_strong_slopes.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(strong_predictions | strong_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_strong_slopes_compare.pdf", height = 9, width = 14)
```

Plot "mild" slopes only
```{r, fig.height=9, fig.width=7, echo=FALSE}
mild_predictions_by_taxon <- iso.sim.model.post.pred %>% filter(taxon %in% c("D", "E", "F")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_manual(values = c("#33a02c","#fb9a99","#e31a1c"), name = "estimated value\nby taxon") +
  scale_color_manual(values = c("#33a02c","#fb9a99","#e31a1c"), name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom", legend.title = element_text(size = 12))
ggsave(mild_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_mild_slopes_by_taxon.png", height = 9, width = 7)

mild_predictions <- iso.sim.model.post.pred %>% filter(taxon %in% c("D", "E", "F")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_blank(), legend.position = "bottom")
ggsave(mild_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data_mild_slopes.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(mild_predictions | mild_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_mild_slopes_compare.pdf", height = 9, width = 14)
```

Plot "weak" slopes only
```{r, fig.height=9, fig.width=7, echo=FALSE}
weak_predictions_by_taxon <- iso.sim.model.post.pred %>% filter(taxon %in% c("G", "H", "I", "J")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, color = taxon, fill = taxon), shape = 16, size = 3, point_interval = "median_qi", .width = 0.89, alpha = 0.6)+
  scale_fill_manual(values = c("#fdbf6f","#ff7f00","#cab2d6","#6a3d9a"), name = "estimated value\nby taxon") +
  scale_color_manual(values = c("#fdbf6f","#ff7f00","#cab2d6","#6a3d9a"), name = "estimated value\nby taxon")+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_text(size = 12), legend.position = "bottom")
ggsave(weak_predictions_by_taxon, filename = "figures/posterior_predictions_varying_slopes_simulated_data_weak_slopes_by_taxon.png", height = 9, width = 7)

weak_predictions <- iso.sim.model.post.pred %>% filter(taxon %in% c("G", "H", "I", "J")) %>%
  ggplot()+
  stat_halfeye(aes(x = value, y = site, shape = "estimated value"), color = "#2C8FE6", fill = "#75B2E6", size = 3, point_interval = "median_qi", .width = 0.89)+
  new_scale(new_aes = "shape")+ 
  geom_star(data = dsim, mapping = aes(x = wc_sim, y = site, starshape = "simulated value"), size = 4, fill = "white", starstroke = 1.5)+
  scale_starshape_manual(values = c("simulated value" = 1), name = " ")+
  xlab("proportion woody cover") + 
  theme_classic(base_size = 16)+
  theme(legend.title = element_blank(), legend.position = "bottom")
ggsave(weak_predictions, filename = "figures/posterior_predictions_varying_slopes_simulated_data_weak_slopes.png", height = 9, width = 7)
```

```{r, fig.height=9, fig.width=14}
(weak_predictions | weak_predictions_by_taxon) + plot_layout(axes = "collect_y")
ggsave("figures/posterior_predictions_varying_slopes_simulated_data_weak_slopes_compare.pdf", height = 9, width = 14)
```

# Model with real data

Based on the model using simulated data, and previous work examining the relationship between d13C and woody cover, we expect that the accuracy of predictions of woody cover will depend on the community composition and which taxa are sampled. 

```{r}
# create new column in dataset for taxonomy used in this analysis
iso_data$taxon_model <- NA
iso_data[iso_data$Family == "Bovidae",]$taxon_model <- iso_data[iso_data$Family == "Bovidae",]$Tribe
iso_data[iso_data$Family != "Bovidae",]$taxon_model <- iso_data[iso_data$Family != "Bovidae",]$Family
```

```{r}
# make pretreatment_detail column into a factor 
iso_data$Pretreatment_Detail <- as.factor(iso_data$Pretreatment_Detail)
levels(iso_data$Pretreatment_Detail) # 7 categories in the detailed method column
```

Because the distributions of woody cover are non-normally distributed, we will model median woody cover at each site (although note that the mean and median values at each site are pretty well correlated, so the result should not be dramatically different):

```{r}
# convert wc_median column to a proportion
iso_data$pct_wc_median <- iso_data$wc_median / 100
```

## Specify the model
```{r}
# model formula - requires a non-linear parameterization because of the two types of varying effects
iso.model.2.formula <- bf(pct_wc_median ~ 0 + d + p,
                          d ~ 1 + d13C_scaled + (1 + d13C_scaled | taxon_model),
                          p ~ 0 + Pretreatment_Detail,
                          nl = TRUE)

# default priors
get_prior(formula = iso.model.2.formula, data = iso_data, family = Beta())

# set the priors 
iso.model.priors <- c(prior("normal(0, 0.5)", class = "b", nlpar = "d"), prior("gamma(15, 2)", class = "phi"), prior("gamma(2,10)", class = "sd", nlpar = "d"), prior("lkj(2)", class = "cor"), prior("normal(0, 0.5)", class = "b", nlpar = "p"))
```

## Prior predictive check
```{r}
# prior predictive check
iso.model.2.pp.check <- brm(
   iso.model.2.formula,
   prior <- iso.model.priors,
   sample_prior = 'only',
   data = iso_data,
   family = Beta(),
   iter = 1000,
   chains = 4, 
   cores = 4,
   file = "models/with_data/iso_model_2_pp_check")

iso.model.2.pp.check
plot(iso.model.2.pp.check, nvariables=3, ask=F)

set.seed(74839)
pp_check(iso.model.2.pp.check , ndraws = 100)
ggsave("figures/prior_predictive_check_iso_model_2.png", bg = "white")
```

## Sample posterior
```{r}
iso.model.2 <- brm(
   iso.model.2.formula,
   prior <- iso.model.priors,
   data = iso_data,
   family = Beta(),
   chains = 4, 
   cores = 4,
   iter = 3000,
   control = list(adapt_delta = 0.99),
   file = "models/with_data/iso_model_2")

iso.model.2
```

```{r}
iso.model.2$fit
```

Posterior predictive check:
```{r}
pp_check(iso.model.2, ndraws = 300)
ggsave("figures/posterior_predictive_check_iso_model_2.png", bg = "white", height = 8, width = 9)
```

## Visualize model output

```{r}
iso.model.2.post <- as_draws_df(iso.model.2) # sample 2000 posterior draws
```

### (Supplementary Figure 4) Plot the chains:
```{r}
iso.model.2.post %>%
  select(-starts_with("r_"), -lp__, -lprior) %>%
  pivot_longer(cols = c(1:13)) %>%
  mutate(`.chain` = as.character(`.chain`)) %>%
  ggplot(aes(x = `.iteration`, y = value, group = `.chain`, color = `.chain`)) +
  scale_color_manual(values = c('#016450', '#3690c0','#a6bddb','#d0d1e6'), name = "chain")+
  geom_line(linewidth = 1/4)+
  ylab(NULL)+
  xlab(NULL) +
  theme_classic()+
  facet_wrap(~name, ncol = 3, scales = "free_y")+
  theme(legend.position = c(.5,.05), legend.direction = "horizontal")
ggsave("figures/trace_plot_pop_effects.png", width= 10 , height = 10)
```

### Fixed effects:
```{r, warning=FALSE, message=FALSE}
# fixed effects minus pretreatment coefficients
iso.model.2.post %>% 
  dplyr::select(b_d_Intercept, b_d_d13C_scaled, sd_taxon_model__d_Intercept, sd_taxon_model__d_d13C_scaled, cor_taxon_model__d_Intercept__d_d13C_scaled) %>%
  pivot_longer(cols = c(b_d_Intercept, b_d_d13C_scaled, sd_taxon_model__d_Intercept, sd_taxon_model__d_d13C_scaled, cor_taxon_model__d_Intercept__d_d13C_scaled)) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for fixed effects (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/fixed_effects_iso_model_2.png", width = 9, height = 7)

# pretreatment effects
iso.model.2.post %>% 
  dplyr::select(b_p_Pretreatment_Detail0.1M_buff_AA:b_p_Pretreatment_DetailNone) %>%
  pivot_longer(b_p_Pretreatment_Detail0.1M_buff_AA:b_p_Pretreatment_DetailNone) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = name), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for pretreatment effects (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/pretreatment_effects_iso_model_2.png", width = 9, height = 7)
```

### Varying intercepts:
(Need to wrangle here to get the coefficients that are the sum of population-level effects and group-level effects, not just the individual group-level effects, which represent a deviance from the corresponding population-level effect).

```{r, warning=FALSE, message=FALSE}
# extract summed model coefficients
varef_model2 <- coef(iso.model.2, summary = F)$taxon_model[,,]

# plot of summed coefficients
varef_model2[,,"d_Intercept"] %>%
  as_data_frame() %>%
  gather() %>%
  mutate(key = fct_reorder(key, desc(value))) %>%
  ggplot() +
  geom_rect(aes(xmin = -1.09, xmax = -0.17, ymax = 24.5, ymin=0), fill = "gray80")+
  geom_vline(xintercept = -0.63, linetype = 2, alpha = 0.5)+
  stat_pointinterval(aes(x =value, y = key), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  ylab(NULL) +
  xlab("coefficients for varying intercepts (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/sum_varying_intercept_coefficients_iso_model_2.png", width = 9, height = 7)
```

### Varying slopes:
```{r, warning=FALSE, message=FALSE}
# plot of summed coefficients
varef_model2[,,"d_d13C_scaled"] %>%
  as_data_frame() %>%
  gather() %>%
  mutate(key = fct_reorder(key, desc(value))) %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = key), color = "#300948")+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  geom_vline(xintercept = -1.02)+
  geom_vline(xintercept = -0.73)+
  geom_vline(xintercept = -0.43)+
  ylab(NULL) +
  xlab("coefficients for varying slopes (log-odds)") +
  theme_classic(base_size = 16)
ggsave("figures/sum_varying_slopes_iso_model_2.png", width = 9, height = 7)
```

### Figure 5: Plot varying effects with family and tribe-level colors for comparison
```{r}
# varying intercepts - family level (minus bovids)
var_int_fam <- varef_model2[,,"d_Intercept"] %>%
  as_data_frame() %>%
  gather() %>%
  filter(key %in% c("Canidae","Cercopithecidae","Elephantidae","Equidae","Felidae","Giraffidae", "Hippopotamidae","Hominidae","Hyaenidae","Hystricidae","Rhinocerotidae","Suidae","Tragulidae","Viverridae")) %>%
  mutate(key = fct_reorder(key, desc(value))) %>%
  ggplot() +
  annotate(geom = "rect", xmin = -1.09, xmax = -0.17, ymax = 14.5, ymin=0, fill = "gray90")+
  geom_vline(xintercept = -0.63, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  stat_pointinterval(aes(x =value, y = key, color = key))+
  family_colors()+
  ylab(NULL) +
  xlim(-3,2)+
  xlab("varying intercepts (log-odds)") +
  theme_classic(base_size = 16)+
  theme(legend.position = "none")
ggsave("figures/sum_varying_intercepts_families.png", width = 9, height = 8)

# varying intercepts - bovid tribes only
var_int_tribe <- varef_model2[,,"d_Intercept"] %>%
  as_data_frame() %>%
  gather() %>%
  filter(key %in% c("Aepycerotini","Alcelaphini","Antilopini","Bovini","Caprini","Cephalophini","Hippotragini","Neotragini","Reduncini","Tragelaphini")) %>%
  mutate(key = fct_reorder(key, desc(value))) %>%
  ggplot() +
  annotate(geom = "rect", xmin = -1.09, xmax = -0.17, ymax = 10.5, ymin=0, fill = "gray90")+
  geom_vline(xintercept = -0.63, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  stat_pointinterval(aes(x =value, y = key, color = key))+
  bovid_colors()+
  ylab(NULL) +
  xlim(-3,2)+
  xlab("varying intercepts (log-odds)") +
  theme_classic(base_size = 16)+
  theme(legend.position = "none")
ggsave("figures/sum_varying_intercepts_tribes.png", width = 9, height = 8)

# varying slopes - family level (minus bovids)
var_slo_fam <- varef_model2[,,"d_d13C_scaled"] %>%
  as_data_frame() %>%
  gather() %>%
  filter(key %in% c("Canidae","Cercopithecidae","Elephantidae","Equidae","Felidae","Giraffidae", "Hippopotamidae","Hominidae","Hyaenidae","Hystricidae","Rhinocerotidae","Suidae","Tragulidae","Viverridae")) %>%
  mutate(key = fct_reorder(key, desc(value))) %>%
  ggplot() +
  annotate(geom = "rect", xmin = -1.02, xmax = -0.43, ymax = 14.5, ymin=0, fill = "gray90")+
  geom_vline(xintercept = -0.73, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  stat_pointinterval(aes(x =value, y = key, color = key))+
  family_colors()+
  ylab(NULL) +
  xlim(-3.3,1)+
  xlab("varying slopes (log-odds)") +
  theme_classic(base_size = 16)+
  theme(legend.position = "none")
ggsave("figures/sum_varying_slopes_families.png", width = 9, height = 8)

# varying slopes - bovid tribes only
var_slo_tribe <- varef_model2[,,"d_d13C_scaled"] %>%
  as_data_frame() %>%
  gather() %>%
  filter(key %in% c("Aepycerotini","Alcelaphini","Antilopini","Bovini","Caprini","Cephalophini","Hippotragini","Neotragini","Reduncini","Tragelaphini")) %>%
  mutate(key = fct_reorder(key, desc(value))) %>%
  ggplot() +
  annotate(geom = "rect", xmin = -1.02, xmax = -0.43, ymax = 10.5, ymin=0, fill = "gray90")+
  geom_vline(xintercept = -0.73, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, color = alpha("gray20", alpha = 0.5)) +
  stat_pointinterval(aes(x =value, y = key, color = key))+
  bovid_colors()+
  ylab(NULL) +
  xlim(-3.3,1)+
  xlab("varying slopes (log-odds)") +
  theme_classic(base_size = 16)+
  theme(legend.position = "none")
ggsave("figures/sum_varying_slopes_tribes.png", width = 9, height = 8)
```

```{r, fig.height=12, fig.width=14}
(var_slo_fam / var_slo_tribe) + plot_layout(axis_titles = "collect", heights = c(14, 10)) | (var_int_fam  / var_int_tribe) + plot_layout(axis_titles = "collect", heights = c(14, 10))
ggsave("figures/summary_varying_effects_color.png", width = 14, height = 12)
```

Extract posterior predictions of woody cover:
```{r, results='hide', message=FALSE}
iso.model.2.post.pred <- posterior_epred(iso.model.2) # generate posterior predictions for each specimen
iso.model.2.post.pred<- t(iso.model.2.post.pred)
iso.model.2.post.pred <- bind_cols(iso_data[,c("locality", "Family", "taxon_model", "Sample_ID", "COUNTRY", "pct_wc_median")], iso.model.2.post.pred)
iso.model.2.post.pred <- pivot_longer(iso.model.2.post.pred, cols = c(7:6006))
```

### (Supplementary Figure 5) Plot overall predictions for each country:
```{r, echo=FALSE}
# uses custom "country_plot" function

post.cameroon.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Cameroon", real_data = iso_data)
ggsave("figures/post_cameroon_2.png", post.cameroon.2, width = 7, height = 4)

post.drc.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "DRC", real_data = iso_data)
ggsave("figures/post_drc_2.png", post.drc.2, width = 8, height = 6)

post.ethiopia.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Ethiopia", real_data = iso_data)
ggsave("figures/post_ethiopia_2.png", post.ethiopia.2, width = 9, height = 8)

post.gabon.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Gabon", real_data = iso_data)
post.gabon.2 <- post.gabon.2 + scale_y_discrete(labels = c("La Lopé")) # fix special character in site name
ggsave("figures/post_gabon_2.png", post.gabon.2, width = 7, height = 4)

post.kenya.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Kenya", real_data = iso_data)
ggsave("figures/post_kenya_2.png", post.kenya.2, width = 10, height = 10)

post.liberia.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Liberia", real_data = iso_data)
ggsave("figures/post_liberia_2.png", post.liberia.2, width = 7, height = 4)

post.moz.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Mozambique", real_data = iso_data)
ggsave("figures/post_mozambique_2.png", post.moz.2, width = 7, height = 4)

post.sa.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "South Africa", real_data = iso_data)
ggsave("figures/post_southafrica_2.png", post.sa.2, width = 9, height = 8)

post.tanzania.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Tanzania", real_data = iso_data)
ggsave("figures/post_tanzania_2.png", post.tanzania.2, width = 8, height = 6)

post.uganda.2 <- country_plot(posterior_preds = iso.model.2.post.pred, country = "Uganda", real_data = iso_data)
ggsave("figures/post_uganda_2.png", post.uganda.2, width = 9, height = 6)
```

```{r, fig.height=20, fig.width=16}
(((post.ethiopia.2 / post.tanzania.2 / post.uganda.2 / post.drc.2 / post.moz.2 / post.cameroon.2 / post.gabon.2 / post.liberia.2) + plot_layout(axis_titles = "collect", heights = c(7,3,4,4,1,1,1,1))) | ((post.kenya.2 / post.sa.2) + plot_layout(axis_titles = "collect", heights = c(21,9)))) + plot_layout(guides = "collect") & theme(legend.position = "bottom")
ggsave("figures/posterior_predictions_summary_all_sites_model_2.pdf", width = 16, height = 20)
```

### (Supplementary Figure 6) Plot by taxon
```{r, echo=F}
post.cameroon.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Cameroon", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_cameroon_2.png", post.cameroon.taxa.2, height = 4, width = 7)

post.drc.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "DRC", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_drc_2.png", post.drc.taxa.2, height = 6, width = 8)

post.ethio.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Ethiopia", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_ethiopia_2.png", post.ethio.taxa.2, height = 8, width = 10)

post.gabon.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Gabon", real_data = iso_data)
post.gabon.taxa.2 <- post.gabon.taxa.2 + scale_y_discrete(labels = c("La Lopé")) # fix special character in site name
ggsave("figures/posterior_predictions_iso_model_by_taxon_gabon_2.png", post.gabon.taxa.2, height = 4, width = 7)

post.kenya.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Kenya", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_kenya_2.png", post.kenya.taxa.2, height = 12, width = 13)

post.liberia.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Liberia", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_liberia_2.png", post.liberia.taxa.2, height = 4, width = 7)

post.moz.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Mozambique", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_mozambique_2.png", post.moz.taxa.2, height = 4, width = 7)

post.sa.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "South Africa", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_sa_2.png", post.sa.taxa.2, height = 8, width = 9)

post.tanzania.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Tanzania", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_tanzania_2.png", post.tanzania.taxa.2, height = 6, width = 8)

post.uganda.taxa.2 <- country_taxon_plot(posterior_preds = iso.model.2.post.pred, country = "Uganda", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_by_taxon_uganda_2.png", post.uganda.taxa.2, width = 9, height = 6)
```

```{r, fig.height=20, fig.width=16}
(((post.ethio.taxa.2 / post.tanzania.taxa.2 / post.uganda.taxa.2 / post.drc.taxa.2 / post.moz.taxa.2 / post.cameroon.taxa.2 / post.gabon.taxa.2 / post.liberia.taxa.2) + plot_layout(axis_titles = "collect", heights = c(7,3,4,4,1,1,1,1))) | ((post.kenya.taxa.2 / post.sa.taxa.2) + plot_layout(axis_titles = "collect", heights = c(21,9)))) + plot_layout(guides = "collect") & theme(legend.position = "bottom", legend.direction = "horizontal")
ggsave("figures/posterior_predictions_summary_all_sites_model_2_taxa.pdf", width = 16, height = 20)
```

### (Supplementary Figure 7) Plot bovids by tribe

```{r}
# can't plot Cameroon or Liberia because there are no bovids in either sample

post.drc.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "DRC", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_drc_2.png", post.drc.bovids.2, height = 6, width = 8)

post.ethio.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "Ethiopia", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_ethiopia_2.png", post.ethio.bovids.2, height = 8, width = 10)

post.gabon.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "Gabon", real_data = iso_data)
post.gabon.bovids.2 <- post.gabon.bovids.2 + scale_y_discrete(labels = c("La Lopé")) # fix special character
ggsave("figures/posterior_predictions_iso_model_bovids_gabon_2.png", post.gabon.bovids.2, height = 6, width = 8)

post.kenya.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "Kenya", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_kenya_2.png", post.kenya.bovids.2, height = 12, width = 13)

post.moz.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "Mozambique", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_mozambique_2.png", post.moz.bovids.2, height = 4, width = 8)

post.sa.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "South Africa", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_sa_2.png", post.sa.bovids.2, height = 8, width = 10)

post.tanzania.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "Tanzania", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_tanzania_2.png", post.tanzania.bovids.2, height = 5, width = 8)

post.uganda.bovids.2 <- country_bovid_plot(posterior_preds = iso.model.2.post.pred, country = "Uganda", real_data = iso_data)
ggsave("figures/posterior_predictions_iso_model_bovids_uganda_2.png", post.uganda.bovids.2, width = 9, height = 6)
```

```{r, fig.height=20, fig.width=16}
(((post.ethio.bovids.2 / post.tanzania.bovids.2 / post.uganda.bovids.2 / post.drc.bovids.2 / post.moz.bovids.2 /  post.gabon.bovids.2 ) + plot_layout(axis_titles = "collect", heights = c(7,2,4,4,1,1))) | ((post.kenya.bovids.2 / post.sa.bovids.2) + plot_layout(axis_titles = "collect", heights = c(21,9)))) + plot_layout(guides = "collect") & theme(legend.position = "bottom", legend.direction = "horizontal")
ggsave("figures/posterior_predictions_summary_all_sites_model_2_bovids.pdf", width = 16, height = 20)
```

### Figure 4: Summary plot of how predictions at all sites compare to real mean woody cover

```{r}
# summary statistics for posterior predictions by site
summary.iso.model.2 <- iso.model.2.post.pred %>% group_by(locality) %>% mutate(mean_wc_pred = mean(value), med_wc_pred = median(value), var_wc_pred = var(value), sd_wc_pred = sd(value), iqr_wc_pred = IQR(value), q1_wc_pred = quantile(value, 0.25), q3_wc_pred = quantile(value, 0.75)) %>% dplyr::select(locality, mean_wc_pred, med_wc_pred, var_wc_pred, sd_wc_pred, iqr_wc_pred, q1_wc_pred, q3_wc_pred) %>% unique()

# join summary statistics with data from woody cover dataset
summary.iso.model.2 <- left_join(summary.iso.model.2, iso_data[,c("locality", "wc_mean", "wc_median", "wc_variance", "wc_stdev", "AREA..sq.km.")], by = "locality") %>% unique()

# convert woody cover mean and median values to proportion
summary.iso.model.2[,c("wc_mean", "wc_median")] <- summary.iso.model.2[,c("wc_mean", "wc_median")]/100
```

```{r}
# plotting median and iqr for posterior predictions
ggplot(data = summary.iso.model.2, aes(y = med_wc_pred, x = wc_median, fill = wc_median)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2)+
  geom_errorbar(aes(ymin=q1_wc_pred, ymax = q3_wc_pred, width=0))+
  geom_point(shape = 21, size = 4)+
  scale_fill_distiller(palette = "BrBG", direction = 1, name = NULL)+
  xlim(0,1)+
  ylim(0,1)+
  labs(y = "posterior predictions", x = "median woody cover")+
  theme_classic(base_size = 16)+
  theme(legend.position = "bottom")+
  guides(fill = guide_colorbar(barwidth = 15))
ggsave("figures/posterior_vs_actual_woody_cover_model_2.png", height = 8, width = 9)

```


## Compare mean/median of each taxon to actual median 
```{r}
# averaged at the site level
taxa_compare <- iso.model.2.post.pred %>% select(-name, -Sample_ID) %>% group_by(locality, taxon_model) %>% mutate(taxon_mean = mean(value), taxon_median = median(value)) %>% select(-value) %>% distinct()
taxa_compare <- taxa_compare %>% mutate(median_diff = pct_wc_median - taxon_median)
taxa_compare <- taxa_compare %>% mutate(abs_med_diff = abs(median_diff))

# family level groupings
taxa_compare %>% filter(Family != "Bovidae") %>% 
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_med_diff)), y = median_diff, color = taxon_model))+
  geom_boxplot()+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_hline(aes(yintercept = -0.1), linetype = 2, color = "gray")+
  geom_hline(aes(yintercept = 0.1), linetype = 2, color = "gray")+
  geom_point(alpha = 0.5)+
  family_colors()+
  labs(y = "actual median - median predicted value (by locality)", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/family_groups_error_boxplots_by_site.png", width = 9, height = 8)

# absolute value
taxa_compare %>% filter(Family != "Bovidae") %>% 
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_med_diff)), y = abs_med_diff, color = taxon_model))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  family_colors()+
  labs(y = "absolute value of\nactual median - median predicted value (by locality)", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/abs_val_family_groups_error_boxplots_by_site.png", width = 9, height = 8)

# bovids
taxa_compare %>% filter(Family == "Bovidae") %>% 
  ggplot(aes(x = fct_reorder(taxon_model, desc(median_diff)), y = median_diff, color = taxon_model))+
  geom_boxplot()+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_hline(aes(yintercept = -0.1), linetype = 2, color = "gray")+
  geom_hline(aes(yintercept = 0.1), linetype = 2, color = "gray")+
  geom_point(alpha = 0.5)+
  bovid_colors()+
  labs(y = "actual median - median predicted value (by locality)", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/tribe_groups_error_boxplots_by_site.png", width = 9, height = 8)

taxa_compare %>% filter(Family == "Bovidae") %>% 
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_med_diff)), y = abs_med_diff, color = taxon_model))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  bovid_colors()+
  labs(y = "absolute value of\nactual median - median predicted value (by locality)", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/abs_val_tribe_groups_error_boxplots_by_site.png", width = 9, height = 8)

```

### Figure 6
```{r}
# using the individual values (taken as the median of the posterior predictions)
individual_compare <- iso.model.2.post.pred %>% select(-name) %>% group_by(Sample_ID) %>% mutate(indiv_mean = mean(value), indiv_median = median(value)) %>% select(-value) %>% distinct()
individual_compare <- individual_compare %>% mutate(diff = pct_wc_median - indiv_median)
individual_compare <- individual_compare %>% mutate(abs_dif = abs(diff))

# plot raw differences for non-bovid families; in descending order based on the absolute value of differences
fam_diff <- individual_compare %>% filter(Family !="Bovidae") %>%
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_dif)), y = diff, color = Family))+
  geom_boxplot()+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_hline(aes(yintercept = -0.1), linetype = 2, color = "gray")+
  geom_hline(aes(yintercept = 0.1), linetype = 2, color = "gray")+
  geom_point(alpha = 0.5)+
  family_colors()+
  ylim(-0.6,0.75)+
  labs(y = "median woody cover - predicted value", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/family_groups_error_boxplots_by_individual.png", width = 9, height = 8)

# plot absolute values of differences for non-bovid families
abs_fam_diff <- individual_compare %>% filter(Family !="Bovidae") %>%
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_dif)), y = abs_dif, color = Family))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  family_colors()+
  labs(y = "absolute value of \nmedian woody cover - predicted value", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/abs_val_family_groups_error_boxplots_by_individual.png", width = 9, height = 8)

# plot raw differences for bovid tribes; in descending order based on the absolute value of differences
bov_diff <- individual_compare %>% filter(Family =="Bovidae") %>%
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_dif)), y = diff, color = taxon_model))+
  geom_boxplot()+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_hline(aes(yintercept = -0.1), linetype = 2, color = "gray")+
  geom_hline(aes(yintercept = 0.1), linetype = 2, color = "gray")+
  geom_point(alpha = 0.5)+
  bovid_colors()+
  ylim(-0.6,0.75)+
  labs(y = "median woody cover - predicted value", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/tribe_groups_error_boxplots_by_individual.png", width = 9, height = 8)

# plot absolute values of differences for bovid tribes
abs_bov_diff <- individual_compare %>% filter(Family =="Bovidae") %>%
  ggplot(aes(x = fct_reorder(taxon_model, desc(abs_dif)), y = abs_dif, color = taxon_model))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  bovid_colors()+
  labs(y = "absolute value of \nmedian woody cover - predicted value", x = NULL)+
  theme_classic(base_size = 16)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("figures/abs_val_tribe_groups_error_boxplots_by_individual.png", width = 9, height = 8)

```

```{r}
(fam_diff | bov_diff) + plot_layout(axis_titles = "collect_y", axes = "collect_y", widths = c(14,10))
ggsave("figures/all_groups_error_boxplots.png", width = 12, height = 8)
```

### Figure 7
Facet plots of each taxon d13C versus distance to locality median and locality woody cover
```{r}
individual_compare <- left_join(individual_compare, iso_data[,c("Sample_ID", "Taxonomic_ID", "d13C_corrected")])

# plot individual d13C vs difference between real and predicted - color based on taxon
individual_compare %>% filter(Family %in% c("Cercopithecidae", "Elephantidae", "Equidae", "Felidae", "Giraffidae", "Hippopotamidae", "Hominidae", "Hyaenidae", "Rhinocerotidae", "Suidae")) %>%
  ggplot(aes(x = d13C_corrected, y = diff, fill = taxon_model))+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_point(shape = 21, alpha = 0.5, size = 3)+
  family_fills()+
  theme_classic(base_size = 16)+
  ylab("median woody cover - predicted value")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model, nrow = 2)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = "none")
ggsave("figures/diff_vs_d13C_family.png", width = 15, height = 8)

# plot individual d13C vs difference between real and predicted - color based on woody cover at locality
individual_compare %>% filter(Family %in% c("Cercopithecidae", "Elephantidae", "Equidae", "Felidae", "Giraffidae", "Hippopotamidae", "Hominidae", "Hyaenidae", "Rhinocerotidae", "Suidae")) %>%
  ggplot(aes(x = d13C_corrected, y = diff, fill = pct_wc_median))+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_point(shape = 21, alpha = 0.6, size = 3)+
  scale_fill_distiller(palette = "BrBG", direction = 1, name = "proportion\nwoody cover")+
  theme_classic(base_size = 16)+
  ylab("median woody cover - predicted value")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model, nrow = 2)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = "bottom", legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 15))
ggsave("figures/diff_vs_d13C_family_wc_colors.png", width = 15, height = 9)

# plot individual d13C vs woody cover at locality
individual_compare %>% filter(Family %in% c("Cercopithecidae", "Elephantidae", "Equidae", "Felidae", "Giraffidae", "Hippopotamidae", "Hominidae", "Hyaenidae", "Rhinocerotidae", "Suidae")) %>%
  ggplot(aes(x = d13C_corrected, y = pct_wc_median, fill = taxon_model))+
  geom_point(shape = 21, alpha = 0.5, size = 3)+
  family_fills()+
  theme_classic(base_size = 16)+
  ylab("median woody cover")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model, nrow = 2)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = "none")
ggsave("figures/wc_vs_d13C_family.png", width = 15, height = 8)

# plot individual d13C (Bovids) vs difference between real and predicted woody cover - color based on tribe
individual_compare %>% filter(Family == "Bovidae" & taxon_model != "Caprini") %>%
  ggplot(aes(x = d13C_corrected, y = diff, fill = taxon_model))+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_point(shape = 21, alpha = 0.5, size = 3)+
  bovid_fills()+
  theme_classic(base_size = 16)+
  ylab("median woody cover - predicted value")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = "none")
ggsave("figures/diff_vs_d13C_bovids.png", width = 9, height = 12)

# plot individual d13C (Bovids) vs difference between real and predicted woody cover - color based on locality woody cover
individual_compare %>% filter(Family == "Bovidae" & taxon_model != "Caprini") %>%
  ggplot(aes(x = d13C_corrected, y = diff, fill = pct_wc_median))+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_point(shape = 21, alpha = 0.5, size = 3)+
  scale_fill_distiller(palette = "BrBG", direction = 1, name = "proportion\nwoody cover")+
  theme_classic(base_size = 16)+
  ylab("median woody cover - predicted value")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = "bottom", legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 15))
ggsave("figures/diff_vs_d13C_bovids_wc_colors.png", width = 9, height = 13)

# plot individual d13C (Bovids) vs woody cover
individual_compare %>% filter(Family == "Bovidae" & taxon_model != "Caprini") %>%
  ggplot(aes(x = d13C_corrected, y = pct_wc_median, fill = taxon_model))+
  geom_point(shape = 21, alpha = 0.5, size = 3)+
  bovid_fills()+
  theme_classic(base_size = 16)+
  ylab("median woody cover")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = "none")
ggsave("figures/wc_vs_d13C_bovids.png", width = 9, height = 12)

# (FIGURE 7) plot ALL groups difference vs d13C with coloring according to locality percent woody cover
individual_compare %>% filter(Family %in% c("Bovidae","Cercopithecidae", "Elephantidae", "Equidae", "Felidae", "Giraffidae", "Hippopotamidae", "Hominidae", "Hyaenidae", "Rhinocerotidae", "Suidae") & taxon_model != "Caprini") %>%
  ggplot(aes(x = d13C_corrected, y = diff, fill = pct_wc_median))+
  geom_hline(aes(yintercept = 0), color = "gray")+
  geom_point(shape = 21, alpha = 0.6, size = 3)+
  scale_fill_distiller(palette = "BrBG", direction = 1, name = "median proportion\nwoody cover")+
  theme_classic(base_size = 18)+
  ylab("median woody cover - predicted value")+
  xlab(expression(paste(delta^{13},"C (\U2030 VPDB)")))+
  facet_wrap(~taxon_model)+
  scale_x_continuous(breaks = c(-20, -10, 0))+
  theme(legend.position = c(0.9,0.1), legend.direction = "horizontal", legend.title.position = "top")+
  guides(fill = guide_colorbar(barwidth = 10))
ggsave("figures/diff_vs_d13C_ALL_wc_colors.png", width = 16, height = 16)
```
